Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-13 arxiv/diff_cov.tex
various edits to polish draft

The different covariates case differs from the same covariates case in two aspects.
First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
Second, the features of different tasks may have different shapes, a phenomenon that is often called ``covariate shift''.
In this section, we show how to deal with these two aspects.

Unlike the previous section, we no longer have an optimal solution for $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies tight generalization bounds for any \emph{local minimum} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Suppose that $X^{(1)} = Z^{(1)}{\Sigma^{(1)}}^{1/2} \in \real^{n_1 \times p}$ and $X^{(2)} = Z^{(2)}{\Sigma^{(2)}}^{1/2} \in \real^{n_2 \times p}$, where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. The matrices $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

Our key result characterizes the asymptotic limit of the inverse of the sum of two different sample covariance matrices. Recall that in the hard parameter sharing, we have to take $r<t$. Hence in the two-task case, we must have $r=1$. 
To motivate our study, consider a special case where $A_1=A_2=1$. 
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \ \ \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.
The empirical loss of $\hat{\beta}_2^{\MTL}$ is close to its expectation similar to Claim \ref{claim_pred_err}.
Hence, we consider the expected loss conditional on the covariates as follows:
\begin{align}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}

Interestingly, the expected loss also admits a nice bias-variance decomposition.
Our main result in this section characterizes the asymptotic limit of the above equation, when both sample size and feature dimension become increasingly large at a fixed ratio.
The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
Intuitively, the scaling of $\hat{\Sigma}^{-1}$ not only depends on both tasks' sample sizes, but also depends on their population covariance matrices.
We introduce a key quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$ that captures the covariate shift between task one and task two.
Let $U\Lambda V^\top$ denote the singular value decomposition of $M$, where $\Lambda\in\real^{n\times p}$ consists of $M$'s singular values $\lambda_1\ge \lambda_2\ge \dots \ge \lambda_p$ as its diagonal entries in descending order.
Our main result is stated as follows.

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be a fixed value in $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) $Z^{(1)}, Z^{(2)}$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits.

	\noindent (i) The variance equation \eqref{eq_variance_2task} satisfies the following estimate with high probability:
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{p^{-1}\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \le  p^{-c_{\varphi}}, %\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) The bias equation \eqref{eq_bias_2task} satisfies the following limit with high probability for any fixed vector $w\in\real^p$ with unit norm,
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \le  p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following self-consistent equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2}, \ \ 
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_2 a_3-a_1 a_4 )}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Our result extends Fact \ref{fact_tr} on the inverse of a single sample covariance matrix.
When $n_1$ is zero, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to equation \eqref{lem_cov_shift_eq}.

\paragraph{How does hard parameter sharing scale with sample size and covariate shift?}
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we illustrate how varying them impact prediction loss, respectively.
\begin{example}[Sample ratio]\label{ex_sample_ratio}
	We first consider the impact of varying sample sizes.
	Consider the random-effects model from Section \ref{sec_same}, with both tasks having an isotropic population covariance matrix.

@@

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) both tasks sample sizes are at least $\frac{3}{2}p$;
	(ii) noise variance is smaller than the shared signal variance: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) task-specific variance is much smaller than the shared signal variance: $d^2 \le p^{-\delta}{\kappa^2}$ for a small constant $\delta>0$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} + p^{-c } \cdot \OO(d^2 + \sigma^2).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
	 with high probability for any fixed $c\in(0, \min(\frac{1}{4}, \delta,\frac{\varphi-4}{2\varphi}))$.
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}

@@

Next, we illustrate an interesting phenomenon where adding task one's samples helps task two initially, but hurts eventually.
Consider the limiting estimate on the left hand side of equation \eqref{cor_MTL_error}.
We vary sample ratio $n_1 / n_2$ by fixing $n_2$ and increasing $n_1$.
%As a function of the sample ratio, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
%We describe a sketch of the proof.
The variance term always reduces as $n_1$ increases.
The bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
\begin{enumerate}
	\item When $\frac{2d^2}{d^2 + \sigma^2} > \frac{p}{n_2 - p}$, the prediction loss decreases monotonically as $n_1$ increases.
	\item When $\frac{2d^2}{d^2 + \sigma^2} < \frac{p}{n_2 - p}$, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
	To see this, near the point where $n_1$ is zero, one can verify that bias increases less than the variance decreases, and there is only one positive root for the derivative being zero.
%Combined together, we observe that the \textit{optimal} sample ratio between $n_1$ and $n_2$ should be among one of following three cases:
%(i) $n_1 = 0$, in which case the prediction loss reduces to $\frac{\sigma^2 p}{n_2 - p}$;
%(ii) $n_1 = \infty$, in which case the prediction loss reduces to $d^2$;
%(iii) the critial point where the derivative of the bias and the variance terms w.r.t. $n_1$ equals zero.
\end{enumerate}

%Based on this observation, we have the following trend when we fix $n_2$ and increase $n_1$:
%\begin{itemize}
%	\item If $d^2$ is smaller than $\frac{\sigma^2 p}{n_2 - p}$, then the prediction loss of hard parameter sharing is always lower than STL. That is, task one always transfers positively to task two.
%	\item Otherwise, task one transfers positively to task two initially, but transfers negatively eventually.
%\end{itemize}
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we observe negative transfer.
%of MTL increases and converges to $pd^2$ as the sample ratio $\rho_1/\rho_2$ increases, and the variance reduction $\frac{\sigma^2}{ \rho_2-1}  -\frac{\sigma^2}{\rho_1+\rho_2-1}$ also increases and converges to $\frac{\sigma^2}{\rho_2-1}$ as $\rho_1/\rho_2$ increases. The tradeoff between these two terms depends on $\rho_1$ in a nonlinear way that is more complicated than the one in the case of varying model distance. But from our results, one can check the following phenomena. If $\frac{\sigma^2}{\rho_2-1} > pd^2$, then the transfer is always positive. On the other hand, if $\frac{\sigma^2}{\rho_2-1} < pd^2$, then there is a transition from the positive transfer to negative transfer as $\rho_1/\rho_2$ increases.
%Therefore, while the variance of MTL still reduces compared to STL,
%If the bias increases more than the amount of variance reduced, we will observe negative transfer.
%Figure \ref{fig_size} provides a simulation result for such a setting (described in Appendix \ref{app_synthetic}).
%We observe that as $n_1 / n_2$ increases, there is a transition from positive to negative transfer.
%There are several unexplained observations near $y = 0$ caused by the concentration error on the right-hand side of \eqref{cor_MTL_error}.

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example studies the effect of varying covariate shifts on the variance limit in equation \eqref{lem_cov_shift_eq}:
%in the left hand side of equation \eqref{lem_cov_shift_eq}:
	\[ \bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} \rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}. \]
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance limit is determined by the spectrum of $M$. To be clear, for this example we assume that the bias is $0$.

>>>>

The previous section assumes that all tasks have the same sample size and feature vectors.
This section studies how different sample sizes and different features impact hard parameter sharing.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
The setting where features differ across tasks is often also called ``covariate shift''.

Unlike the previous section, we no longer have an optimal solution for $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies tight generalization bounds for any \emph{local minimum} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Suppose that $X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p}$ and $X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p}$, where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. The matrices $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

Our key result characterizes the asymptotic limit of the inverse of the sum of two different sample covariance matrices.
To motivate our study, consider a special case where $A_1=A_2=1$. 
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By Proposition 1 of \citet{WZR20}, we consider the case of $r = 1 < t = 2$, since when $r > 1$, the optimal solution of $f(A, B)$ is equivalent to single-task learning.
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
Hence we can write down a closed form equation for any local minimum of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \ \ \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to its expectation similar to Claim \ref{claim_pred_err}):
\begin{align}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Our main result in this section characterizes the asymptotic limit of the above equation under the high-dimensional assumption.
Intuitively, the spectrum of $\hat{\Sigma}^{-1}$ (and hence its trace) not only depends on both tasks' sample sizes, but also depends on the ``alignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
We introduce a key quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and as we show below, the trace of $\hat{\Sigma}^{-1}$ has an intricate dependence on the spectrum of $M$.

Let $U\Lambda V^\top$ denote the SVD of $M$ and let $\lambda_1, \lambda_2, \dots, \lambda_p$ denote $M$'s singular values in descending order.
Our main result is stated as follows.

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be a fixed value in $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have the following limits.

	\noindent (i) The variance equation \eqref{eq_variance_2task} satisfies the following estimate with high probability:
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{p^{-1}\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \le  p^{-c_{\varphi}}, %\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) The bias equation \eqref{eq_bias_2task} satisfies the following limit with high probability for any fixed vector $w\in\real^p$ with unit norm,
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \le  p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following self-consistent equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2}, \ \ 
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_2 a_3-a_1 a_4 )}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Part (i) of our result extends Fact \ref{fact_tr} to the inverse of the sum of two sample covariance matrices.
To see this, when $n_1$ is zero, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to equation \eqref{lem_cov_shift_eq}.

\paragraph{How does hard parameter sharing scale with sample size and covariate shift?}
One can see that the bias-variance limits depend intricately on both tasks' samples sizes and covariate shift.
Next, we illustrate how varying them impact prediction loss, respectively.
\begin{example}[Sample size ratio]\label{ex_sample_ratio}
	We first consider the impact of varying sample sizes.
	Consider the random-effects model from Section \ref{sec_same}, with both tasks having an isotropic population covariance matrix.

@@

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) both tasks sample sizes are at least $\frac{3}{2}p$;
	(ii) noise variance is smaller than the shared signal variance: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) task-specific variance is much smaller than the shared signal variance: $d^2 \le p^{-\e}{\kappa^2}$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} +  \OO(p^{-c/2}).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
%	 with high probability for any fixed $c\in(0, \min(\frac{1}{4}, \delta,\frac{\varphi-4}{2\varphi}))$.
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}

@@

Next, we illustrate an interesting phenomenon where adding task one's samples helps task two initially, but hurts eventually.
Consider the limiting estimate on the left hand side of equation \eqref{cor_MTL_error}.
We vary sample ratio $n_1 / n_2$ by fixing $n_2$ and increasing $n_1$.
%As a function of the sample ratio, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
%We describe a sketch of the proof.
The variance term always reduces as $n_1$ increases.
The bias term always increases as $n_1$ increases, which can be verified by calculating the derivative of the bias term.
Furthermore, by calculating the derivative of the bias-variance limits with respect to $n_1$, we obtain the following dichotomy (details omitted).
\begin{enumerate}
	\item When $\frac{2d^2}{d^2 + \sigma^2} < \frac{p}{n_2 - p}$, the prediction loss decreases monotonically as $n_1$ increases.
	\item When $\frac{2d^2}{d^2 + \sigma^2} > \frac{p}{n_2 - p}$, the limiting estimate always decreases first from $\frac{\sigma^2 p}{n_2 - p}$ with $n_1$ being zero, and then increases to $d^2$ when $n_1$ goes to infinity.
	To see this, near the point where $n_1$ is zero, one can verify that bias increases less than the variance decreases, and there is only one positive root for the derivative being zero.
\end{enumerate}

\begin{example}[Covariate shift]\label{ex_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.
Our second example focuses on how varying covariate shifts impacts the \textit{variance} limit in equation \eqref{lem_cov_shift_eq}, which shows that
%in the left hand side of equation \eqref{lem_cov_shift_eq}:
	\[ \bigtr{\Sigma^{(2)} \hat{\Sigma}^{-1}} \rightarrow \frac{1}{n_1 + n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}= \frac{1}{n_1 + n_2} \bigtr{(a_1 M^{\top} M + a_2 \id)^{-1}}. \]
%As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
Hence the variance limit is determined by the spectrum of $M$.% To be clear, for this example we assume that the bias is $0$.

To illustrate the above result, suppose that half of $M$'s singular values are equal to $\lambda > 1$ and the other half are equal to $\lambda^{-1}$.
In particular, when $\lambda = 1$, there is no covariate shift.
As $\lambda$ increases, the severity of covariate shift increases.
We observe the following dichotomy.
\begin{itemize}
	\item If $n_1 \ge n_2$, then the variance limit is smallest when there is no covariate shift.
	\item If $n_1 < n_2$, then the variance limit is largest when there is no covariate shift.
\end{itemize}
\end{example}
We explain why the dichotomy happens. The variance estimate for this example is equal to
\begin{align*}
	\frac{p}{2(n_1 + n_2)} f(\lambda), \ \ \text{ where }\ \  f(\lambda) = {(\lambda^{-2} a_1 + a_2)^{-1} + (\lambda^2 a_1 + a_2)^{-1}}.
\end{align*}
Using the fact that $a_1 + a_2 = 1 - \frac{p}{n_1 + n_2 }$, we can obtain that
\begin{align*}
	f(\lambda) - f(1) &= \left(\lambda^2 a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1}
	+ \left(\lambda^{-2} a_1 + \frac{n_1 + n_2 - p}{n_1 + n_2} - a_1\right)^{-1}  - \frac{2(n_1 + n_2)}{n_1 + n_2 - p} \\
	&= \left(2a_1 - \frac{n_1 + n_2-p} {n_1 + n_2 }\right)  g(\lambda, a_1), %\cdot (\lambda^2-1)^2
\end{align*}
where $g(\lambda, a_1) > 0$ is a fixed function and can be derived from algebraic calculations (details omitted).
We now show that $a_1 \ge \frac{n_1 + n_2-p}{2(n_1 + n_2 )}$ if and only if $n_1 \ge n_2$, and hence explain the dichonomy.
In fact, if $a_1>a_2$, then the equations in \eqref{eq_a12extra} give that $a_1> \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$, and one can verify that
$$ 
 \frac{n_1}{n_1 + n_2}=a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} 
> a_1 + \frac{p}{2(n_1+n_2)} \left(\frac{\lambda^2}{\lambda^2+1}+\frac{\lambda^{-2}}{\lambda^{-2}+1}\right) =\frac{1}{2}.
$$
This implies $n_1>n_2$. Similarly, if $a_1<a_2$, equations in \eqref{eq_a12extra} give that $a_1 < \frac{n_1 + n_2-p}{2 (n_1 + n_2)}$ and $n_1<n_2$. Thus, we conclude that $f(\lambda) \ge f(1)$ if and only if $n_1 \ge n_2$.

\paragraph{Overview of Stieltjes transform.}
For the rest of this section, we present an overview of the proof of Theorem \ref{thm_main_RMT}.
The central quantity of interest is the inverse of the sum of two sample covariance matrices $\hat{\Sigma}^{-1}$.
We note that the variance equation $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}]$ is equal to $(n_1 + n_2)^{-1} \bigtr{W^{-1}}$, where $W$ is
\begin{align}\label{eigen2extra}
	W \define \frac{1}{n_1 + n_2}(\Lambda U^\top (Z^{(1)})^\top Z^{(1)} U\Lambda  + V^\top (Z^{(2)})^\top Z^{(2)}V).
\end{align}
This formulation is helpful because we know that ${Z^{(1)}}^{\top} Z^{(1)}$ and ${Z^{(2)}}^{\top} Z^{(2)}$ are both sample covariance matrices with isotropic population covariance, and $U, V$ are both orthonormal matrices.

Our proof uses the Stieltjes transform or the resolvent method in random matrix theory.
We briefly describe the key ideas and refer interested readers to classical texts such as  \citet{bai2009spectral,tao2012topics,erdos2017dynamical}.
For any probability measure $\mu$ supported on $[0,\infty)$, the Stieltjes transform of $\mu$ is given by
$$m_\mu(z):= \int_0^\infty \frac{\dd\mu(x)}{x-z}, \ \ \text{ for any complex number } z\in \C\setminus [0,\infty).$$
The Stieltjes transform method reduces the study of a probability measure $\mu$ to the study of a complex function $m_\mu(z)$.
%To study the trace of $\hat{\Sigma}^{-1}$, we consider the Stieltjes transform of its empirical spectral distribution.
